Parameteric controls on responses to biogeochemical forcing in the CLM5

### 2018-11-15 main.tex

\authors{Rosie Fisher\affil{1},
Will Wieder\affil{1},
Ben Sanderson\affil{1},
Charlie Koven\affil{2},
Keith Oleson\affil{1},
Chonggang Xu\affil{3},
Ashehad Ali\affil{1},
Josh Fisher\affil{4},
Mingjie Shi\affil{4},
Katie Dagon\affil{1},
Danica Lombardozzi\affil{1},
Anthony Walker\affil{5},
Gordon Bonan\affil{1},
David Lawrence\affil{1}
}

\title{Nitrogen cycling in the CLM5: controls on responses to environmental forcing}
\author{rosie fisher}
\date{April 2018}

@@

\affiliation{2}{Lawrence Berkeley National Laboratory, Berkeley, California, USA}

\affiliation{3}{Los Alamos National Laboratory, Los Alamos, New Mexico, USA}

\affiliation{4}{NASA Jet Propulsion Laboratory, Pasadena, California, USA}

\affiliation{5}{Oak Ridge National Laboratory, Oak Ridge, Tennessee, USA}

@@

\item Here we assess the sensitivity of the model to a suite of parameters pertinent to the cycle of Carbon and Nitrogen. We assess the dependence of the both the model state, and also the responses of the system to higher CO2, Nitrogen and temperature to parameter uncertainty. Model runs are conducted at representative individual sites in temperate, tropical and boreal systems.

\item The model responses to CO2 and N fertilization are primarily linked to the representation of the fraction of plants that can fix Nitrogen, and to the costs of N in the environment.

\item Identification of the primary factors driving ecosystem responses to fertilization both assists in the interpretation of global simulations and in the identification of priorities for fieldwork and process updates.

\item This implies that a critical control in the CLM, and in coupled simulations with the CESM for CMIP are dependant on the representation of Nitrogen fixations and its dynamics.

\end{keypoints}

@@

Simulating the cycling of nutrients, and how nutrient availability impacts ecosystem growth and function, has been repeatedly identified as a crucial element of Earth system models (\cite{piao2013}, \cite{gruber2008}, \cite{wang2009}). In the CMIP5 multi-model intercomparison process, only one model, the Community Land Model v4.0, included an active nitrogen (N) cycle. The CLM4.0 projected that inclusion of N dynamics might significantly limit the ability of the terrestrial biosphere to respond to fertilization by increasing atmospheric carbon dioxide (\cite{friedlingstein2006},\cite{friedlingstein2014}, \cite{arora2013}). Subsequent and parallel to CMIP5, many terrestrial biosphere and land surface models that represent N dynamics using alternative structural assumptions have been developed \cite{wang2007}, \cite{zaehle2010}, \cite{goll2012}, \cite{smith2014}). Nonetheless, substantial uncertainty remains in the land surface modeling community regarding how nutrient limitations should be represented in models (\cite{zaehledalmonech2011}).

Unlike representations of photosynthetic responses to environmental conditions, or radiative transfer through forest canopies, our understanding of how nutrient cycling functions both within plants and in whole ecosystems is not governed by any well-tested and widely accepted paradigms. Detailed model-data synthesis activities, notably those conducted at North American free-air carbon dioxide enrichment (FACE) experiments, have revealed important differences in model behavior emerging in part from the absence of a broadly accepted theoretical framework for construction and parameterization of nutrient cycling processes (\cite{zaehle2014}). Few, if any studies, however, have looked at how CO$_{2}$ or N fertilization are governed by the parameters or structure of a given model.

The Community Land Model is the land surface representation within the Community Earth System Model (CESM, \cite{hurrell2013}). Version 5 of the CLM was recently released to the research community as part of the CESM2.0 release. An overview of model developments is presented by \cite{lawrence2018}, a descriptions of the global Nitrogen cycling by Wieder et al. (in prep), the crop model (Lombardozzi et al. in prep), the land use components (Lawrence P et al. in prep), land surface hydrology (Swenson et al. in prep) and plant hydraulics (Kennedy et al. in prep).

The CLM5 model includes numerous major update of the nitrogen cycling representation subsequent to CLM4.0 and CLM4.5. The introduction of several new dynamical features into the CLM5 has implications for both parameteric uncertainty and responses to environmental forcing. Wieder et al. (in prep) report the global-scale implications of the full model update and note that the CLM( responds more strongly to CO2 and less strongly to N fertilization than its predecessors. In this paper, we describe the new components of the biogeochemical cycling of nitrogen, their impacts upon the model system state and responses to environmental forcing.

@@

The CLM5 nitrogen cycling builds on the developments of the implementation of the CLM4.0 'CN' model (\cite{thornton2007}), and on modifications to the biogeochemical cycling (notably vertically stratified soil decomposition processes) included in the interim CLM4.5 model (\cite{koven2013}, \cite{bonan2012}). The new model integrates three major additional prognostic elements of the Nitrogen cycle, including: \\

@@

2) The `FUN' module (Fixation and Uptake of Nitrogen), which simulates the dynamics of Nitrogen acquisition from the environment (\cite{fisher2010fun}, \cite{brzostek2014}, \cite{shi2016}).\\

3) The `FLEXCN' N cycle implementation, which is used here primarily to allow variation in tissue C:N ratios, and is adapted from \cite{ghimire2016}.\\

@@

\subsubsection{LUNA}

Most land surface models predict photosynthetic parameters from leaf nitrogen content (\cite{kattge2009}, \cite{bonan2012}). There is significant evidence, however, that plant photosynthetic capacity can respond to environmental conditions, including CO$_{2}$ fertilization (\cite{ainsworth2007}) and changing temperature (\cite{hikosaka2005}), irradiance (\cite{niinemets1998}) and soil moisture (\cite{keenan2009}). Given the importance of photosynthetic capacity as a key model parameter (\cite{rogers2017}) the use of a static photosynthetic capacity appears to be a poor means of capturing its variation in time and space and responses to climatic change (\cite{walker2017}).

The LUNA model predicts the optimal balance of of $V_{c,max}$ and $J_{max}$ for the prevailing time-averaged environmental conditions. The model determines the rate constants that are consistent with a co-limitation of photosynthesis by both electron capture and carboxylation processes. In this fashion, $V_{c,max}$ is primarily controlled by the N per unit leaf area ($N_{area}$, which is itself a function of the target C/N ratio, the specific leaf area, and the prognostic variation in leaf C/N ratios) and environmental conditions. The capacity of the LUNA model to represent appropriate responses to temperature and CO$_{2}$ was assessed by \cite{xu2012} while the global calibration and the geographical predictions of the model were described by \cite{ali2016}.

@@

One major issue confronting N cycle models is how to deal with this 'excess' carbon under N limiting conditions (\cite{zaehle2010}, \cite{dekauwe2014}). The FUN model operates on the principle that for the alternative sources of Nitrogen acquisition by plants (active uptake from soil, retranslocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon. Further, the cost of each of these uptake pathways is variable in both time and space. FUN hypothesizes that plants will take up Nitrogen from pathways which are 'cheaper' to them. Where N is scarce in the environment, carbon that would otherwise have been used for plant growth can be deployed to acquire (pay for) more N. The central element of FUN is a simultaneous equation which assumes, firstly, that ($C_{avail}$), the total carbon available (after maintenance respiration and replenishment of stores are accounted for) is either used for plant growth ($C_{growth}$) or for Nitrogen uptake, ($C_{nuptake}$).

@@

Further, the Nitrogen acquired from the environment ($N_{uptake}$) must equal that which is deployed in the growing plant tissues. The carbon used for growth is further reduced by the growth respiration term ($f_{gr}$) which applies only to tissues that are constructed:

\begin{equation}
N_{uptake}=C_{nuptake}/N_{cost} =\frac{C_{growth}}{(CN_{target}/(1.0+f_{gr})}
\end{equation}

Thus the Carbon that is expended on N uptake is determined by:

@@

The average Nitrogen cost ($N_{cost}$) is derived from the simultaneous uptake of N across all uptake streams, while the target CN ratio ($CN_{target}$) is the combined target CN ratio of all the plant tissues, weighted for the size of the different pools. The FUN model is documented \href{https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/FUN/CLM50_Tech_Note_FUN.html}{here}

A further issue addressed by FUN is the preferential use of symbiotic N fixation when soil mineral N concentrations are low (\cite{vitousek2002}). The CLM4.0 and CLM4.5 predict N fixation as a linear function of net primary productivity (\cite{wieder2015} illustrate the impacts of uncertainty in this function) but FUN considers fixation rates to be an emergent property of plant assimilation rates (providing carbon for fixation), the relative costs of environmental N acquisition, and temperature (which exerts a primary control over the enzymatic processes, following \cite{houlton2008}). Fixation is preferred when the costs of N acquisition are higher than that of fixation. Note that while the CLM5 has no representation of biological succession (\cite{fisher2018}), this property nonetheless affects the amount of time taken to achieve a spun-up state from bare ground, in productive areas, N fixation is preferred and soil N builds up much quicker. In low productivity areas, carbon is required to fix N from a 'col start' initial state, impacting the survival of marginal plants; thus higher initial stocks of N are required to allow growth once metabolic and turnover costs have been met in early spin-up.

\subsubsection{FLEXCN}

Several elements of model behaviour that are critical for simulation of plant responses to variation in nitrogen availability emerged from the FACE model data synthesis activity that were not captured by previous versions of the CLM Nitrogen cycle (\cite{zaehle2014}, \cite{medlyn2015using}). For example, changes in the C:N ratio accounted for large changes in ecosystem carbon storage, but these are not captured by the static tissue C:N ratios used in the CLM4.0 and CLM4.5 nitrogen cycle. For this reason, we introduce a flexible tissue C:N ratio into the CLM5.

\cite{ghimire2016} implemented a suite of modifications to the CLM Nitrogen cycling model, including variable tissue C:N ratio, prognostic $V_{c,max}$ (resulting from this dynamic leaf N content) and also an alternative Nitrogen uptake algorithm. In the default configuration of CLM5, the LUNA model predicts $V_{c,max}$ from leaf N content (while also taking into account environmental conditions), and the FUN model represents Nitrogen uptake, the primary element of CLM5 utilized from the \cite{ghimire2016} model is the modification to the allocation scheme that allows tissue C:N ratios to be modified in response to varying supply of Nitrogen (although the alternative schemes implemented by \cite{ghimire2016} are available as options). In this new allocation algorithm, the total Nitrogen supply in each timestep is partitioned between tissues according to their relative 'demand' terms (ascertained from the target C:N ratio and the carbon allocated to the pool) and C:N ratios vary with Nitrogen supply rates. This new allocation code we term the FLEXCN module.

@@

In this paper, we assess in detail the sensitivity of the CLM5 model structure to its input parameters. We use\\

1. A range of single site experiments \\

2. One-at-a-time and global parameter uncertainty analyses\\

3. CO$_{2}$ fertilization experiments and\\

4. Nitrogen addition experiments.\\

to fully investigate the performance of the model under steady-state and step-change modifications in forcing. Given the large amount of analysis and attention devoted to assessing the impacts of transient CO$_{2}$ and N in the context of global feedback loop analysis, we thus consider it a worthwhile exercise to illustrate how parametric choices in the CLM5 might affect overall behaviour in larger-scale experiments, including the controls on global carbon fertilization `$\beta$' values. We also hope to highlight areas of process and parameter uncertainty that appear to require greater attention in future model development and testing cycles.

Because of the increased number of prognostic processes with the CLM5, many alternative representation of C:N cycle interactions are possible within the same model structural space, including an effective removal of N fixation capacity (where carbon available for fixation is fixed at 0$\%$ of assimilation) and variation in the degree to which to C:N ratios are flexible. We further investigate the consequences of turning off the LUNA model and returning to the fixed leaf N/$V_{c,max}$ relationship from \cite{ghimire2016}.

@@

For this experiment we utilized six different experimental sites. Two in tropical forests (Caxiuana and Barro Colorado Island) two in temperate forest (the Oak Ridge National Lab - Free Air Carbon Dioxide Enrichment (FACE) experiment and Harvard Forest), one in high altitude forest (Niwot Ridge, Colorado) and one in seasonally dry evergreen forest (Metolius Forest).

\subsection{Simulation Protocol}

We spun up the default version of the model for 400 years at each site, in 'accelerated spin-up' mode (where the turnover of slower cycling carbon and nitrogen pools is accelerated, see (/cite{lawrence}) using pre-industrial CO$_{2}$ concentrations (274ppm) and recycling the available site-level meteorological inputs. 400 years was sufficient to bring all carbon and nitrogen pools into equilibrium. Subsequent to this, parameters were perturbed (see below), and a second spin-up conducted for 180 years at each site, with accelerated mode turned off. Using the state at the end of the perturbed parameter spin-ups, we ran transient simulations for each site and parameter combination starting in 1850 and invoking transient CO$_{2}$ and Nitrogen deposition until the end of the observed meteorological forcing. We ran additional CO$_{2}$ and Nitrogen deposition experiments by increasing the CO$_{2}$ concentration to 550ppm. For the Oak Ridge FACE site, we only increased CO$_{2}$ in the growing season, in line with the experimental protocol applied there and using the observed CO$_{2}$ time-series. For all other sites, we increased CO$_{2}$ to 550ppm year round starting in 1997**. For Nitrogen, we added 5kgN $m^{-2}$/year, also starting in 1997, so that the results can be compared with those reported by Wieder et al. (in prep).

\subsection{Perturbed Physics Ensemble (PPE)}

The CLM5 has a very large number of parameters to which the C and N cycles are potentially sensitive. Here we focused on parameters of particular relevance to the C and N cycles and on those which are newly introduced by the modified N cycle. We choose a larger set of parameters for one-at-a-time analyses, followed by a smaller set of these to which we subject the model to global sensitivity tests. Knowledge of model structure as well as previous investigations of the CLM model informs our choice of parameters. Our principle goal is to identify parameters that might have large impacts on responses to model forcing, as opposed to changes in the base state of the model, since these are the primary determinants of the carbon cycle feedback in the context of the Earth System Model. However, we also report responses of the model state to perturbations. Additional investigations into the parametric sensitivity of the hydrology and surface energy balance in the CLM5 are planned (Dagon et al. in prep) and into the plant hydraulics parameterizations (Kennedy et al in prep), thus, we have not included the full range of those parameters here.

In the first instance, we conduct one-at-a-time (OAAT) perturbations of the parameters around a mean state, to allow a straightforward visualization and interpretation of the results. For each parameter, for each site, we conduct four simulations, with two above and two below the default value (as well as the default simulation).

@@

The parameters included in our sensitivity analysis (Table \ref{table_parameters}) are: \\

1) \textbf{Specific Leaf Area}, \emph{`SLA'}, which is a critical parameter determining the area of canopy (in m$^{2}$) derived from one gram of leaf biomass. This parameter is known to be one of the primary elements of the leaf economics spectrum (\cite{wright2004}), and is well-characterized by the TRY database (\cite{kattge2011}).\\

2) \textbf{Fine root area per unit leaf area}, \emph{`froot\_leaf'}, the ratio of fine root biomass to leaf area. Unlike in previous versions of the CLM, fine root biomass now affects the capacity of the model to acquire both water and nutrients (via the resistance to water flow and the cost of N acquisition), thus the impact of this parameter is of greater interest than it's previous functionality as a simple 'tax' on plant growth. \\

3) \textbf{Stem to leaf area ratio}, \emph{`stem\_leaf'}, reflects the balance between accumulation of stem and woody biomass, and leaf area. Increases in this parameter decrease the leaf area index achievable for a given productivity, and conversely increase the equilibrium woody biomass predictions.\\

4) \textbf{Nitrogen Costs}, \emph{`n\_costs'}, are a set of parameters that determine the environmental cost of Nitrogen uptake from the soils. These six parameters were previously defined by \cite{brzostek2014}. Here and in CLM5 in general, we maintain the ratios of the parameters as defined by the existing FUN model calibration, but then allow the magnitudes of the parameters to change concomitantly, since the definitions of soil N availability vary between CLM5 and \cite{brzostek2014}. These parameters include \emph{ekc\_active} and \emph{ekn\_active}, the rate constants that relate the cost of ectomycorhhizal uptake to root carbon (ekc) and soil N concentration (ekn), and their counterparts \emph{akc\_active} and \emph{akn\_active} for arbuscular mycorrhizal fungi, and \emph{kc\_nonmyc} and \emph{kn\_nonmyc} for non-mycorrhizal active uptake. \\

5) \textbf{The fraction of carbon that can be used for fixation}, \emph{`'fracfixers'}. This parameter is the fraction of the assimilation (gross primary productivity - autotropihic respiration, or $gpp$ - $R_{auto}$) that can be used for symbiotic N fixation by the FUN model, and is a proxy for the ecosystem-level fractional productivity of Nitrogen fixing plants, in lieu of, for example, a prognostic model of N fixer distributions, or an input map of their abundance, both of which could plausibly be added in future generations of CLM. \\

6) \textbf{The fraction of carbon spent on respiration per unit of new tissue (growth respiration)}, \emph{grperc}, was revised substantially in CLM5 based on Atkin O. (pers comm.) from 0.3 to 0.11. Here we test the impacts of this change on overall carbon budgets.\\

@@

8) \textbf{Slope of stomatal conductance model}, \emph{medlyn\_slope}. This new parameter, which replaces the fixed Ball-Berry slope used in CLM5, (and is PFT-dependant), is critically linked to the response of ecosystems to CO$_{2}$, assessed in this set of experiments. \\

>>>>

\authors{Rosie Fisher\affil{1},
William Wieder\affil{1},
Benjamin Sanderson\affil{2},
Charles Koven\affil{3},
Keith Oleson\affil{1},
Chonggang Xu\affil{4},
Ashehad Ali\affil{1},
Joshua Fisher\affil{5},
Mingjie Shi\affil{5},
Anthony Walker\affil{6},
Gordon Bonan\affil{1},
David Lawrence\affil{1}
}

\title{Parameteric controls on responses to biogeochemical forcing in the CLM5}
\author{Rosie Fisher}
\date{November 2018}

@@

\affiliation{2}{CERFACS, Toulouse, France}

\affiliation{3}{Lawrence Berkeley National Laboratory, Berkeley, California, USA}

\affiliation{4}{Los Alamos National Laboratory, Los Alamos, New Mexico, USA}

\affiliation{5}{NASA Jet Propulsion Laboratory, Pasadena, California, USA}

\affiliation{6}{Oak Ridge National Laboratory, Oak Ridge, Tennessee, USA}

@@

\item Here we assess the sensitivity of the model to a suite of parameters pertinent to the cycling of Carbon and Nitrogen. We assess the dependence of the both the model state, and also the responses of the system to CO$_{2}$ and Nitrogen fertilization to parametric uncertainty. Model runs are conducted at representative individual sites in temperate, tropical and boreal systems.

\item The model responses to CO$_{2}$  and N fertilization are primarily linked to the representation of the fraction of plants that can fix Nitrogen, the costs of N in the environment and stoichiometric flexibility of plant tissues, all of which are subject to large empirical and theoretical uncertainty.

\item Identification of the primary factors driving ecosystem responses to fertilization both assists in the interpretation of global simulations and in the identification of priorities for fieldwork and process updates.
\end{keypoints}

@@

Simulating the cycling of nutrients, and how nutrient availability impacts ecosystem growth and function, has been repeatedly identified as a crucial element of Earth system models (\cite{piao2013}, \cite{gruber2008}, \cite{wang2009}). In the CMIP5 multi-model inter-comparison process, only one model, the Community Land Model v4.0, included an active nitrogen (N) cycle. The CLM4.0 projected that inclusion of N dynamics might significantly limit the ability of the terrestrial biosphere to respond to fertilization by increasing atmospheric carbon dioxide (\cite{friedlingstein2006}, \cite{friedlingstein2014}, \cite{arora2013}). Subsequent and parallel to CMIP5, many terrestrial biosphere and land surface models that represent global N dynamics using alternative structural assumptions have been developed (\cite{wang2007}, \cite{zaehle2010}, \cite{goll2012}, \cite{smith2014}). Nonetheless, substantial uncertainty remains in the land surface modeling community regarding how nutrient limitations should be represented in models (\cite{zaehle2011}).

Unlike representations of photosynthetic responses to environmental conditions, or radiative transfer through forest canopies, our understanding of how nutrient cycling functions both within plants and in whole ecosystems is not governed by any well-tested and widely accepted paradigms. Detailed model-data synthesis activities, notably those conducted at North American free-air carbon dioxide enrichment (FACE) experiments, have revealed important differences in model behavior emerging in part from the absence of a broadly accepted theoretical framework for construction and parameterization of nutrient cycling processes (\cite{kauwe2013}, \cite{zaehle2014}, \cite{medlyn2015}). Few studies, however, have looked at how CO$_{2}$ or N fertilization are governed by the parameters or structure of a given model  (c.f. \cite{meyerholt2018}).

The Community Land Model is the land surface representation within the Community Earth System Model (CESM, \cite{hurrell2013}). Version 5 of the CLM was recently released to the research community as part of the CESM2.0 release and open-source releases and development are housed at \url{https://github.com/ESCOMP/ctsm/}. An overview of model developments is presented by \cite{lawrence2018}, a description of the global Nitrogen cycling by Wieder et al. (in prep), the crop model by(Lombardozzi et al. in prep), the land use components by (Lawrence P et al. in prep), land surface hydrology by (Swenson et al. in prep) and plant hydraulics by (Kennedy et al. in prep).

CLM5 includes numerous major updates of the nitrogen cycle representation relative to CLM4.0 and CLM4.5. The introduction of several new dynamical features into the CLM5 has implications for both parameteric uncertainty and responses to environmental forcing. Wieder et al. (in prep) report the global-scale implications of the full model update and note that the CLM5 responds more strongly to CO$_{2}$ and less strongly to N fertilization than its predecessors.  In this paper, we describe the new components of the biogeochemical cycling, and investigate the features of the new model that control the model state and biogeochemical forcing responses in depth, using single-site experiments and parametric sensitivity analyses.

@@

The CLM5 nitrogen cycling builds on the implementation of carbon-nitrogen interactions in the CLM4.0(CN) model (\cite{thornton2007}), and on modifications to the biogeochemical cycling (notably vertically stratified soil decomposition processes) included in the CLM4.5 (\cite{koven2013}, \cite{bonan2012}). The new model integrates three major additional prognostic elements of the Nitrogen cycle, including: \\

@@

2) The `FUN' module (Fixation and Uptake of Nitrogen), which simulates the dynamics and carbon economics of N acquisition from the environment (\cite{fisher2010fun}, \cite{brzostek2014}, \cite{shi2016}), and\\

3) The `FLEXCN' N cycle implementation, adapted from \cite{ghimire2016}, which is used in CLM5 primarily to allow variation in tissue C:N ratios,  .\\

@@

\subsubsection{LUNA}
Most land surface models predict photosynthetic parameters from leaf nitrogen content (\cite{kattge2009}, \cite{bonan2012}). There is significant evidence, however, that plant photosynthetic capacity can respond to environmental conditions, including CO$_{2}$ fertilization (\cite{ainsworth2007}) and changing temperature (\cite{hikosaka2005}), irradiance (\cite{niinemets1998}) and soil moisture (\cite{keenan2009}). Given the importance of photosynthetic capacity as a key model parameter (\cite{rogers2017}) the use of a static photosynthetic capacity appears to be a poor means of capturing its variation in time and space and responses to climatic change (\cite{walker2017}).

The LUNA model predicts the optimal balance of of $V_{c,max}$ and $J_{max}$ for the prevailing time-averaged environmental conditions. The model determines the rate constants that are consistent with a co-limitation of photosynthesis by both electron capture and carboxylation processes. In this fashion, $V_{c,max}$ is primarily controlled by the N per unit leaf area $N_{area}$, (which is itself a function of the target C/N ratio, the specific leaf area, and the prognostic variation in leaf C/N ratios) and prevailing environmental conditions (CO$_{2}$, temperature, humidity, soil moisture). The capacity of the LUNA model to represent appropriate responses to temperature and CO$_{2}$ was assessed by \cite{xu2012} while the global calibration and the geographical predictions of the model were described by \cite{ali2016}.

@@

One major issue confronting N cycle models is how to deal with this 'excess' carbon under N limiting conditions (\cite{zaehle2010}, \cite{dekauwe2014}). The FUN model operates on the principle that for the alternative sources of Nitrogen acquisition by plants (active uptake from soil, retranslocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon. Further, the cost of each of each uptake pathway is variable in both time and space. FUN hypothesizes that plants will take up Nitrogen from pathways which are 'cheaper' to them. Where N is scarce in the environment, carbon that would otherwise have been used for plant growth can be deployed to acquire (pay for) more N. Thus, N limitation leads to reduced growth and higher expenditure of C on N uptake (and reduces carbon use efficiency).  The central element of FUN is a set of simultaneous equations which assume, firstly, that the total carbon available (after maintenance respiration and replenishment of stores are accounted for), `$C_{avail}$', is either used for plant growth ($C_{growth}$) or for Nitrogen uptake, ($C_{nuptake}$).

@@

Further, the Nitrogen acquired from the environment ($N_{uptake}$) must equal the amount spent on N divided by it's average cost ($N_{cost}$), and also the amount of carbon deployed for growth divided by the plant CN ratio:

\begin{equation}
N_{uptake}=C_{nuptake}/N_{cost} =\frac{C_{growth}}{(CN_{plant}/(1.0+f_{gr})}
\end{equation}

The carbon used for growth is further reduced by the growth respiration term ($f_{gr}$, which applies only to tissues that are constructed). Rearrangement of these constraints gives a predictive term for the amount of carbon that is expended on N uptake, as:

@@

The average Nitrogen cost ($N_{cost}$) is derived from the uptake of N, weighted for it's cost across all uptake streams, while the plant CN ratio ($CN_{plant}$) is the combined target CN ratio of all the plant tissues, weighted for the size of the different pools. The FUN model is documented at \url{https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/FUN/CLM50_Tech_Note_FUN.html}

One issue addressed by FUN is the preferential use of symbiotic N fixation when soil mineral N concentrations are low (\cite{vitousek2002}). The CLM4.0 and CLM4.5 predict N fixation as a linear function of net primary productivity ((\cite{wieder2015} illustrated the impacts of uncertainty in this function).  FUN considers fixation rates to be an emergent property of plant assimilation rates, the relative costs of environmental N acquisition compmared to fixation, and temperature (which exerts a primary control over the enzymatic processes, following \cite{houlton2008}). Fixation is preferred when the costs of `active' N acquisition from soil are higher than that of fixation. Note that while the CLM5 has no representation of biological succession (\cite{fisher2018}), this property nonetheless affects the amount of time taken to achieve a spun-up state from bare ground. In most cases, N fixation is preferred when soil N is low, and so and soil N builds up much quicker. In low productivity areas, however, carbon is required to fix N from a 'cold start' initial state, impacting the carbon use efficiency and survival of marginal plants; thus higher initial stocks of N are required to allow growth once metabolic and turnover costs have been met in early spin-up.

\subsubsection{FLEXCN}
The results of the FACE model data comparison exercise illustrated that changes in the C:N ratio accounted for large changes in ecosystem carbon storage under elevated CO$_{2}$ (\cite{zaehle2014}, \cite{medlyn2015using}).  These changes were not captured by the static tissue C:N ratios used in the CLM4.0 and CLM4.5 nitrogen cycle. For this reason, we introduce a flexible tissue C:N ratio into the CLM5.

\cite{ghimire2016} implemented a suite of modifications to the CLM Nitrogen cycling model, including variable tissue C:N ratio, prognostic $V_{c,max}$ (resulting from this dynamic leaf N content) and also an alternative Nitrogen uptake algorithm. In the default configuration of CLM5, the LUNA model predicts $V_{c,max}$ from leaf N content (while also taking into account environmental conditions), and the FUN model represents Nitrogen uptake, thus the primary element of CLM5 utilized from the \cite{ghimire2016} model is the modification to the allocation scheme that allows tissue C:N ratios to be modified in response to varying supply of Nitrogen. In the new allocation algorithm, the total Nitrogen supply in each timestep is partitioned between tissues according to their relative 'demand' terms (ascertained from the target C:N ratio and the carbon allocated to the pool) and C:N ratios vary with Nitrogen supply rates.  The allocation scheme retains a fixed-fraction approach to allocation in general, and thus is capable of generating physiologically sub-optimal leaf area index values, for example, with some of the canopy in negative carbon balance, an issue which must be taken into account during simulations under high productivity conditions.

@@

In this paper, we assess in detail the sensitivity of the CLM5 model structure to its input parameters. We use a range of single site experiments, combined with one-at-a-time parameter uncertainty analyses from the default state, as well as CO$_{2}$ fertilization experiments and Nitrogen addition experiments, 
to fully investigate the performance of the model under steady-state and step-change modifications in forcing.

Given the large amount of analysis and attention devoted to assessing the impacts of transient CO$_{2}$ and N in the context of global feedback loop analysis, we thus consider it a worthwhile exercise to illustrate how parametric choices in the CLM5 might affect overall behaviour in larger-scale experiments, including the controls on global carbon fertilization `$\beta$' values. We also hope to highlight areas of process and parameter uncertainty that appear to require greater attention in future model development and testing cycles.

Because of the increased number of prognostic processes with the CLM5, many alternative representation of C:N cycle interactions are possible within the same model structural space, including an effective removal of N fixation capacity (where carbon available for fixation is fixed at 0$\%$ of assimilation) and variation in the degree to which to C:N ratios are flexible. We thus consider that an exploration of the CLM5 parameter space is also equivalent to an exploration of a large part of the relevant model structural space.

@@

For this experiment we utilized six different experimental sites. Two in tropical forests (Caxiuana and Barro Colorado Island) two in temperate forest (the Oak Ridge National Lab - Free Air Carbon Dioxide Enrichment (FACE) experiment and Harvard Forest), one in high altitude forest (Niwot Ridge, Colorado) and one in seasonally dry evergreen forest (Metolius Forest) **tbd**.

\subsection{Simulation Protocol}
We spun up the default version of the model for 400 years at each site, in 'accelerated spin-up' mode (whereby the turnover of slower cycling carbon and nitrogen pools is accelerated, see (\cite{lawrence2018}) using pre-industrial CO$_{2}$ concentrations (274ppm) and recycling the available site-level half-hourly meteorological inputs. 400 years was sufficient to bring all carbon and nitrogen pools into equilibrium. Subsequent to this, parameters were perturbed (see below), and a second spin-up conducted for 180 years for each site, parameter and value combination, with accelerated mode turned off. Restarting from the state at the end of the perturbed parameter spin-ups, we ran 'control' transient simulations for each site and parameter combination starting in 1760 and invoking historical transient CO$_{2}$ and Nitrogen deposition until the end of the observed meteorological forcing.

For each site/parameter/value combination, we also ran one CO$_{2}$ and one Nitrogen deposition experiment. To allow comparison with \cite{wieder}, we increased CO$_{2}$ to 550ppm, starting in 1998. For Nitrogen, we added 5kgN $m^{-2}$/year, also starting in 1998. In line with \cite{wieder}, we assessed the impact of changes after 15 years of fertilization in both cases.

\subsection{Perturbed Physics Ensemble (PPE)}
The CLM5 has a large array of parameters that might have direct and indirect impacts on biogeochemical cycling in the CLM5. Here, we are focused on the recent developments in the vegetation model, and how they might impact upon the responses to CO$_{2}$ and N fertilization. Therefore, 
we focus on parameters of particular relevance to the C and N cycles and on those which are newly introduced by the modified N cycle. The choice of parameters used here is informed primarily by the model structure, as well as many iterative and informal investigations of the CLM5 model during the development process.

Our principle goal is to identify parameters that might have large impacts on responses to model forcing, as opposed to changes in the base state of the model, since these are the primary determinants of the carbon cycle feedback in the context of the Earth System Model. However, we also report responses of the model state to perturbations. Additional investigations into the parametric sensitivity of the hydrology and surface energy balance in the CLM5 are planned (Dagon et al. in prep) and into the plant hydraulics parameterizations (Kennedy et al in prep), thus, we have not included the full range of those parameters here. We expect that further focused analysed will be conducted by the CLM5 community, in tandem with analyses of coupled model experiments.

To reduce the complexity involved in understanding and attributing changes in model behaviour with parameter variation, we conducted one-at-a-time (OAAT) perturbations of the parameters around a mean state, which allows more straightforward visualization and interpretation of the results. For each parameter, for each site, we conduct four simulations, with two above and two below the default value (as well as the default simulation). We recognize that a global parameter sensitivity analysis in this realm would also be of interest, and indeed have conducted many such analyses, but consider that assessment of the output of a global as well as a OAAT analysis of the entire domain of results presented here would reduce the opportunity for understanding the model parameter space in depth, as is the goal of our study.

@@

The parameters included in our sensitivity analysis are both detailed in (Table \ref{table_parameters}), and described in more depth below. \\

1) \textbf{Specific Leaf Area}, \emph{`SLA'}, is a critical parameter, common to almost all land surface models, which determines the area of canopy (in m$^{2}$) derived from one gram of leaf biomass. High SLA values correspond to thinner, 'cheaper' leaves, which are typically associated with shorter leaf lifespan and lower nitrogen per unit leaf area (\cite{wright2004}).\\

2) \textbf{Fine root area per unit leaf area}, \emph{`froot\_leaf'}, is the ratio of fine root biomass to leaf area. Unlike in previous versions of the CLM, the fine root biomass now affects the capacity of the model to acquire both water and nutrients (via the resistance to water flow and the cost of N acquisition), thus the impact of this parameter is of greater interest than it's previous functionality as a simple 'tax' on plant growth. \\

3) \textbf{Stem to leaf area ratio}, \emph{`stem\_leaf'}, controls the allocation of biomass to stem and woody biomass, which in CLM5 is a constant fraction of leaf allocation. Increases in this parameter decrease the leaf area index achievable for a given productivity, and conversely increase the equilibrium woody biomass predictions. Woody biomass in CLM5 has no function and thus there is no cost to plants associated with low values of this parameter\\

4) \textbf{Nitrogen Costs}, \emph{`n\_costs'}, are a set of parameters that determine the environmental cost of Nitrogen uptake from the soils. These six parameters were defined and calibrated by \cite{brzostek2014}. In CLM5 we maintain the ratios of the parameters as defined by the existing FUN model calibration, but then allow the magnitudes of the parameters to change concomitantly, since the definitions of soil N availability vary between CLM5 and \cite{brzostek2014}. The \emph(`n\_costs'}) parameters include \emph{ekc\_active} and \emph{ekn\_active}, the rate constants that relate the cost of ectomycorhhizal uptake to root carbon (ekc) and soil N concentration (ekn), and their counterparts \emph{akc\_active} and \emph{akn\_active} for arbuscular mycorrhizal fungi, and \emph{kc\_nonmyc} and \emph{kn\_nonmyc} for non-mycorrhizal active uptake. \\

5) \textbf{The fraction of carbon that can be used for fixation}, \emph{`fracfixers'}. This parameter is the fraction of the assimilation (gross primary productivity - autotropihic respiration, or $gpp$ - $R_{auto}$) that can be used for symbiotic N fixation by the FUN model, and is a proxy for the ecosystem-level fractional productivity of Nitrogen fixing plants, in lieu of, for example, a prognostic model of N fixer distributions, or an input map of their abundance, both of which could plausibly be added in future generations of CLM. \\

6) \textbf{The fraction of carbon spent on respiration per unit of new tissue (growth respiration)}, \emph{`grperc'}, was revised substantially in CLM5 based on methodolicaly issues with the observation of growth respiration from Atkin O. (pers comm.) from 0.3 to 0.11. CLM5 also adopts a new leaf maintenance respiration scheme, based on \cite{atkin2015}, which significnatly increases maintenance respiration costs. Here we test the impacts of this change on overall carbon budgets.\\

@@

8) \textbf{Slope of stomatal conductance model}, \emph{medlyn\_slope}. This new parameter, which replaces the fixed Ball-Berry slope used in CLM5, (and is PFT-dependant), is critically linked to the response of ecosystems to CO$_{2}$, since it determines the degree of stomatal opening for a given combination of assimilation capacity, CO$_{2}$ and vapour pressure deficit (assuming no soil moisture lmitations).\\
